The Multi-Stencil Framework: A Component-Based Approach to Orchestrate Stencils

### 2017-03-05 ARTICLES/IJPP/articleR1/msl.tex
passe Sec. 4

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
        edge node {} (Du3)
    (DDSp1) edge node {} (DDS)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (Cu)
    (Datap2) edge node {} (Data)
    (Datap1) edge node {} (Data)
    (Data) edge node {} (Datau)
          edge node {} (mpidata)
    (DDSp2) edge node {} (DDS);
\end{tikzpicture}
\vspace*{.5em}
\caption{Generic Assembly according to the Multi-Stencil program formalism.}
\label{fig:ga}
\end{center}
\end{figure}

@@

\paragraph{\textbf{Time}}
As indicated in Figure~\ref{fig:ga}, this component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}). Thus it is composed of a time loop and eventually of a convergence reduction.

\paragraph{\textbf{DDS}}
This component is responsible for the mesh and its entities $\mathcal{M}$ and $\Phi$, the computation domains $\mathcal{D}$, and the set of neighborhood functions $\mathcal{N}$. This component is generic, thus not typed. When GA is instantiated and specialized by MSC, a specific type of mesh is replaced in this DDS meta-component \CP{pas clair}. However, needed interfaces of such a DDS component are well known, and any component providing these interfaces can be indifferently used by MSF: a third party can propose new DDS components. In this paper, both data and task parallelism are used. In the case of data parallelism this component needs to handle mesh partitioning and needs to provide a synchronization interface as it is detailed in Section~\ref{sect:parallelism}. One can note that this component is red, this means that this component should be provided by a HPC specialist in MSF as illustrated in Figure~\ref{fig:msf}.

\paragraph{\textbf{Data}}
This component is inseparable from the previous one $DDS$. It actually uses the DDS component to instantiate one mesh variable. Thus, it is responsible for the set of mesh variables $\Delta$. This component is also red. As $DDS$ it should be provided by a HPC specialist. One can note that this component is represented by a double line box. This means that it can be instantiated multiple time while others are instantiated once. Actually, this component is instantiated as many times as the number of quantities in the simulation.

\paragraph{\textbf{Computations}}
Finally, this component is responsible for $\Gamma$, \ie the computations of the simulation. This component is green and blue because it is partly generated by MSC and partly feeded by the developer. Actually, this component is replaced by a sub-assembly of component, produced by MSC, for which the parallel part is automatically generated. In this generated sub-assembly, components responsible for numerical codes is completed by the developer. This sub-assembly generation is described in Section~\ref{sect:parallelism}.

@@

The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}. A Multi-Stencil program is composed of eight parts as in Equation~(\ref{eq:msp}).

@@

\item The \texttt{mesh} keyword (Fig.~\ref{fig:grammar},~l.1) introduces an identifier for $\mathcal{M}$, the single mesh of the simulation. For example \texttt{cart} in Fig.~\ref{fig:exmsl},~l.1. The language, based on the meta-model is independent of the mesh topology, thus this identifier is actually not used by the compiler but is syntactic sugar for the user.

>>>>

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
        edge node {} (Du3)
    (DDSp1) edge node {} (DDS)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (Cu)
    (Datap2) edge node {} (Data)
    (Datap1) edge node {} (Data)
    (Data) edge node {} (Datau)
          edge node {} (mpidata)
    (DDSp2) edge node {} (DDS);
\end{tikzpicture}
\vspace*{.5em}
\caption{Generic Assembly according to the Multi-Stencil program formalism.
Components circled by a double line identify those that will be instantiated multiple times at execution.
Component colors represent the actor of Figure~\ref{fig:msf} responsible for the component implementation: green for those automatically generated by the compiler, red for those implemented by HPC specialists and blue for those implemented by the developer.}
\label{fig:ga}
\end{center}
\end{figure}

@@

\paragraph{\textbf{Time}}
This component is responsible for the time $T$ defined in Equation~(\ref{eq:msp}). It is composed of a time loop and potentially of a convergence reduction.

\paragraph{\textbf{DDS}}
This component is responsible for the mesh and its entities $\mathcal{M}$ and $\Phi$, the computation domains $\mathcal{D}$, and the set of neighborhood functions $\mathcal{N}$.
This component is generic; its type is a parameter of the generic assembly.
When the generic assembly is instantiated and specialized by MSC, a component handling a specific type of mesh is used in place of the DDS placeholder.
\CP{pas clair}
The interfaces exposed by this component are however well defined and any component providing these interfaces can be indifferently used.
Third party developers can therefore propose new DDS components.
In this paper, both data and task parallelism are used. In the case of data parallelism this component handles mesh partitioning and provides a synchronization interface as detailed in Section~\ref{sect:parallelism}.
The implementation of this component is the responsibility of HPC specialists.

\paragraph{\textbf{Data}}
This component is responsible for the set of mesh variables $\Delta$.
Each instance of the component uses the DDS component to handle one single mesh variable.
It is closely related to $DDS$ and its implementation is typically provided by the same HPC specialist as $DDS$.

\paragraph{\textbf{Computations}}
This component is responsible for $\Gamma$, \ie the computations of the simulation. 
It is automatically replaced by a sub-assembly of components produced by MSC for which the parallel part is automatically generated and the part responsible for the numerical code is filled by the developer.
The sub-assembly generation is described in Section~\ref{sect:parallelism}.

@@

The grammar of the Multi-Stencil Language is given in Figure~\ref{fig:grammar} and an example is provided in Figure~\ref{fig:exmsl}. A Multi-Stencil program is composed of eight parts that match those of Equation~(\ref{eq:msp}).

@@

\item The \texttt{mesh} keyword (Fig.~\ref{fig:grammar},~l.1) introduces an identifier for $\mathcal{M}$, the single mesh of the simulation. For example \texttt{cart} in Fig.~\ref{fig:exmsl},~l.1. The language, based on the meta-model is independent of the mesh topology, thus this identifier is actually not used by the compiler but is syntactic sugar for the user.
\JB{J'aime pas trop l'utilisation de ``syntactic sugar'' ici... Syntactic sugar est generalement utilise pour un racourci de syntaxe (i.e. int x = myInteger au lieu de int x = myInteger.getIntVal()), ce n'est pas le cas ici}
